### 2022-03-04 analysis/paper/paper.Rmd
Added more text to the discussion section

---
title: "Late Pleistocene stone artefact technology at Mau A, Yen Bai, Norther Vietnam"
author:
  - Quinn Habedank:
      email: fl@oneoeurhg.edu
      institute: [UofO]
      correspondence: true
  - Ben Marwick:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

# Abstract

Flaked stone artefacts from Southeast Asia typically lack visually distinctive and strongly patterned forms, which can make them challenging to analyze and interpret. As a result, many of the cultural dynamics of Pleistocene hunter-gatherers of this region are poorly understood. We use 2D shape data to hypothesize a relationship between unretouched flake shape and assemblage reduction intensity at Mau A, an early Holocene archaeological site in northern Vietnam. We apply a Principal Components Analysis to the flake outlines to investigate shape variation throughout the reduction sequence (measured by dorsal cortex coverage). We find that flake shape varies by reduction stage, primarily through differences in flake length and width. Our results suggest that flake shape is sensitive to assemblage reduction intensity, and may give useful comparative insights when other attributes show little variation. These results are important for understanding stone artefact assemblages from Southeast Asia which often yield little variation when analysed with traditional approaches.

## Introduction

- Stone tools from South East Asia tend to lack distinctive typological categories (Mijares 2008; Borel et al. 2013; Borel et al. 2017)
- The purpose of this study was to examine the relationship between flake shape and reduction intensity
- Approach used in the paper is a response to the lack of traditional typological categories in South East Asia
-Further context from (McAdams et al, 2020)

Here is a citation [@Marwick2017]

<!-- BM: In this section we need to review previous work on investigating flake shape and identifying what affects flake shape, not just in Southeast Asia, but anywhere in the world -->

## Excavations at Mau A

Mau A is an open air archaeological site located on the banks above 5 m above the Red River in Yen Bai Province, Northern Vietnam, at the confluence of a small stream entering the river. Excavations were conducted in 2015 by a collaboration including researchers from the University of Washington, the Yen Bai Provincial Museum, the Institute of Archaeology in Hanoi, and the Vietnam National University - Social Sciences and Humanities University. An area of 2 x 2 m was excavated in 13 units of ten centimeters deep, through dense silty clay deposits to a depth of about 1.3 m below the surface. Subtle changes in stratigraphy indicated a slightly sloping deposit with four layers, consistent with reports of excavations in this location in the 1980s. A small amount of plastic and modern ceramics were found in the uppermost 0.1 m of deposit. The most striking feature identified during our excavations was the third layer, which is distinctive as very dense layer of flaked stone artefacts at about 0.6-1.1 m below the surface. No other features were identified during excavation.

This dense lithic layer contained the majority of stone artefacts recovered from the site. The  chaîne opératoire of Mau A lithic technology is composed of unifacial shaping on long cobbles to produce sumatraloids (sumatralith-like pieces). The chaîne opératoire also involved the shaping of thick ovoid cobbles for the production of choppers or chopping-tools and half-cobbles (longitudinally split) that are shaped into tools with transverse cutting edges. Flake scar surfaces and edges are fresh and unweathered, indicating _in situ_ production and limited post-depositional movement. While previous work has claimed the Mau A assemblage is Son Vian (a precurser to Hoabinhian technology, characterised by unifacial flaking of blocky, cubic cobbles)

<!-- BM: need to cite The Oxford Handbook of the Archaeology and Anthropology of Hunter-Gatherers, Ha Van Tan 1997, and the Proceedings of the Prehistoric Society -->

Four radiocarbon ages were obtained from isolated charcoal fragments in the deposit. The charcoal was sampled to bracket the upper and lower boundaries of the dense lithic layer. <!-- add calibrated age numbers here --> The ages indicate a rapid accumulation process for this layer at the terminal Pleistocene.

<!-- BM: add a paragraph on MH's pollen analysis to indicate local palaeoenvironment, and add her as co-author-->

<!-- QH: Trying to think of a way to make this figure work with the current figure numbering system we have in bookdown. Adding in a number that starts before 1.1 for now -->
## Methods

![Figure 1.0. Schematic figure showing both the intial measurements that were used to generate our landmarks and the location of our landmarks on a hypothetical flake. A, B, C, D, and E are the top horizontal measurement, width at a quarter of the maximum length, maximum dimension length, width at a quarter of the maximum length, and width at three quarter of the maximum length respectively. Points 1 through 9 are all of the landmarks generated from measurements A through E.  ](./test_graphic.png)

Dorsal cortex percentages discretized upon standard cut points were used to create reduction categories. Flakes with no dorsal cortex were categorized as tertiary flakes, those with up to a dorsal cortex percentage of 50% were considered secondary and those with a dorsal cortex value greater than 50% were considered primary (Bradbury & Carr 1995).

Mass clusters were generated using univariate k-means, with the end result being 6 mass clusters, with mass cluster 1 having the least mass on average and 6 having the greatest mass on average. Univariate k-means is a non-hierarchical clustering technique that creates a set number of groups equal to k, with points being assigned to groups in such a manner that the total distances between each point and mean of the group is minimized. We used the Ckmeans.1d.dp package in R to perform our univariate k-means analysis (Wang & Song, 2011). The Ckmeans.1d.dp has been used for univariate clustering in applications outside of archaeology (Song & Zhong, 2020).

We initially recorded conventional linear dimension measurements, such as max dimension length and width at various points along the max dimension length, from all flakes recovered from the excavation. Then we converted these linear dimensions into landmarks points to represent the shape of each artefact. The specific linear measurements used to create the 2D landmarks were the top horizontal measurement, the max dimension length, width at a quarter of the maximum length, width at half of the maximum length, and width at three quarters of the maximum length.

<!-- BM: Because our approach here is very novel, most readers will struggle to understand it. To help them, I think we will need a table that defines each of our dimensions, and a figure that shows an example of our landmark points on a typical artefact -->

In analysing the landmarks we followed a conventional shape analysis workflow of Generalized Procrustes Analysis (GPA) followed by Principal Components Analysis (PCA) (Riede & Pedersen, 2018; Hoggard et al., 2019; Radinović & Kajtez, 2021; Okumura & Araujo 2014; Archer et al., 2021; Theska et al. 2020). We used GPA to normalize landmarks for size, orientation, and rotation, leaving only shape data. The normalized landmarks were then analysed by PCA, a form of exploratory data analysis. Principal Component Analysis takes the initial set of data and transforms and reduces them into a smaller set of variables, whist still preserving most of the variation. To explore the output we made PCA biplots of the first two Principal Components to visualize shape variation based upon secondary categorical variables, namely flake reduction categories, excavation units, and mass clusters.

## Results

<!-- BM: basic stats on raw materials and change in technology by depth to satisfy readers about the basic details of the assemblage,

<!-- BM: then QH's more complex analysis on flakes after the basic data -->

<!-- BM: then previous more complex analysis on cores after QH's analysis of flakes -->

## PCA Analysis on Dorsal Cortex Groups

```{r}
library(tidyverse)

# import the data
ma_flakes <- readxl::read_excel(here::here("analysis/data/raw_data/UW Fieldschool Mau A 2015 post_excavation_season_April_2016.xlsx"),
                                col_names = FALSE)

# transpose so that columns are variables, rows are individual artefacts
# and set col data types so we can compute on them
ma_flakes_tr <-
ma_flakes %>%
  pivot_longer(-`...1`)  %>%
  pivot_wider(names_from = `...1`,
              values_from = value) %>%
  mutate_all(parse_guess) %>%
  # fix a few bad guesses
  mutate(across(c(mass,
                thickness_at_0.25_width,
                thickness_at_0.75_width,
                dorsal_cortex_perc,
                dorsal_scar_initiation_count),
         parse_number))

# export this data frame to use with our analysis
write_csv(ma_flakes_tr,
          here::here("analysis/data/derived_data/ma_flakes_tr.csv"))

```

```{r}
# here we import the CSV file of caliper measurements, then compute on the
# measurements to generate coordinates for a shape outline

library(tidyverse)

# import the data 
ma_flakes_tr <- read_csv(here::here("analysis/data/derived_data/ma_flakes_tr.csv"))

# custom function is from here, we have a local version in 999-functions.R
# so we can work offline if we need to
# https://gist.github.com/benmarwick/fac5754131c1fcc14d32b1658215f0e3

source(here::here("analysis/code/999-functions.R"))

polys <- polygon_shape_by_variables(data = ma_flakes_tr,       # our data frame
                                    x = 'width_at_0.50_length',                # col for coord for x-axis of poly centers
                                    y = 'max_dimension',       # col for coord for y-axis of poly centers

                                    sf = 10,    # scale factor
                                    top_horizontal_measurement =    'platform_width', # cols to use to compute poly verts
                                    main_vertical_measurement =     'percussion_length',
                                    upper_horizontal_measurement =  'width_at_0.25_length',
                                    middle_horizontal_measurement = 'width_at_0.50_length',
                                    bottom_horizontal_measurement = 'width_at_0.75_length')

# make ready to convert to TPS format (thin plate spline)
polys_lmk <-
polys %>%
  group_by(fill) %>%
  mutate(landmark.number = row_number(fill),
         specimen.ids  = fill) %>%
  ungroup() %>%
  select(-fill)

# export as TPS file
create_tps(polys_lmk,
           here::here("analysis/data/derived_data/landmarks.tps"))

```

```{r}

library(tidyverse)
library(geomorph)
library(Morpho)
library(mclust)
library(cluster)
library(factoextra)

lands <- readland.tps(
  here::here("analysis/data/derived_data/landmarks.tps"),
  specID = "ID",
  warnmsg = FALSE)

# drop outliers  identified in the previous script
outlier_ids <- c(1152,  606, 1147, 1125,  756 ,1023,  457,
                 813,  415,  877,  677, 1055,  239,  811,  601,  861,
                  794,  103 , 598 ,1110 , 906 ,  59 ,   4)

lands_no_outliers <- lands[, , -outlier_ids]
lands_no_outliers_names <- dimnames(lands_no_outliers)[[3]]

sup.geo <- gpagen(lands_no_outliers,  
                  ProcD=F, 
                  Proj = T,
                  print.progress = FALSE)

# Principal component analysis (PCA) using a base R function
pca.geo <- prcomp(two.d.array(sup.geo$coord))

# TODO: find a meaningful way to break this up into groups
# pass these groups onto the PCA for drawing ellipses

# this block of code below shows one way to discretize cortex percentage
# into three groups. We need to explore other cutoffs, and other grouping
# variables besides dorsal cortex. This code below is a good template for
# starting these explorations
```

```{r, fig.cap="PCA biplot for flake shape and dorsal cortex class."}

ma_flakes_tr_with_groups <-
  ma_flakes_tr %>%
  slice(-c(1152,  606, 1147, 1125,  756 ,1023,  457,
                 813,  415,  877,  677, 1055,  239,  811,  601,  861,
                  794,  103 , 598 ,1110 , 906 ,  59 ,   4)) %>% 
 # filter(!is.na(dorsal_cortex_perc)) %>% 
  mutate(dorsal_cortext_group =
           case_when(
             between(dorsal_cortex_perc, 0, 0) ~ "Tertiary",
             between(dorsal_cortex_perc, 1, 50) ~ "Secondary",
             between(dorsal_cortex_perc, 51, 100) ~ "Primary"
           ))

library(Ckmeans.1d.dp)

test_result_mass <- Ckmeans.1d.dp(ma_flakes_tr_with_groups$mass)

ma_flakes_tr_with_groups <-
  ma_flakes_tr_with_groups %>% 
  mutate(mass_group = as.factor(test_result_mass$cluster))

# ggplot2-based plot of the PCA results using a factoextra function
cortex_biplot <- fviz_pca_ind(pca.geo,
             geom.ind = "point", # show points without labels
            #  line below sets the grouping column for colouring the points
             habillage = ma_flakes_tr_with_groups$dorsal_cortext_group,
             palette = "npg", #color palette, adjust to preferences
             addEllipses = T, #adds ellipses or convex hulls
             ellipse.type = "convex", #convex hulls
             mean.point = F, #don’t show average point
             legend.title = "Dorsal Cortex Class") +  #adjust accordingly)
             coord_fixed(ratio = 1)+ #adjustable for variation-weighted plots
  theme(plot.title = element_blank())

```

```{r, fig.cap="PCA biplot for flake shape and excavation unit."}
# Excavation unit PCA
# ggplot2-based plot of the PCA results using a factoextra function
excavation_unit_biplot <- fviz_pca_ind(pca.geo,
             geom.ind = "point", # show points without labels
            #  line below sets the grouping column for colouring the points
             habillage = as.factor(ma_flakes_tr_with_groups$Unit),
             palette = viridis::cividis(13), # color palette, adjust to preferences
             addEllipses = T, #adds ellipses or convex hulls
             ellipse.type = "convex", #convex hulls
             mean.point = F, #don’t show average point
             legend.title = "Excavation\nunit",
             title = "Change over time in flake shape") +  #adjust accordingly)
             coord_fixed(ratio = 1)  + #adjustable for variation-weighted plots
             theme(panel.background = element_rect(fill='white',
                                        colour = "white"),
              plot.background = element_rect(fill='white',
                                       colour = "white")) +
    theme(plot.title = element_blank())
```

<!-- This portion is lifted largely from the ISKM presentation. Might well reword it later, but it will do for now -->

```{r, fig.cap="PCA biplot for flake shape and mass cluster groups"}

# Mass PCA
# ggplot2-based plot of the PCA results using a factoextra function
mass_biplot <- fviz_pca_ind(pca.geo,
             geom.ind = "point", # show points without labels
            #  line below sets the grouping column for colouring the points
             habillage = ma_flakes_tr_with_groups$mass_group,
             palette = viridis::viridis(6), #color palette, adjust to preferences
             addEllipses = T, #adds ellipses or convex hulls
             ellipse.type = "convex", #convex hulls
             mean.point = F, #don’t show average point
             legend.title = "Mass\ncluster",  #adjust accordingly
            title = "Flake mass and flake shape") + 
             coord_fixed(ratio = 1) + #adjustable for variation-weighted plots
             theme(panel.background = element_rect(fill='white',
                                        colour = "white"),
              plot.background = element_rect(fill='white',
                                       colour = "white")) +
    theme(plot.title = element_blank())
```

<!-- Below are my attempts to create a nice looking plot_grid of the three biplots. I was not able to get it to display properly, so I am just printing off each biplot individually all in a row as a temporary measure.  -->
```{r}
library(cowplot)

# plot_grid(
#  cortex_biplot, NULL, NULL,  excavation_unit_biplot,  mass_biplot, NULL,
#  ncol = 2,
#  labels = c("A", NULL, NULL,"B","C",NULL),
#  label_size = 12,
#  align = "v",
#  rel_heights = c(1.1,
#                  1,
#                  1.1,
#                  4,
#                  1,
#                  1),
#  rel_widths = c(2,
#                 2,
#                 2,
#                 4,
#                 2,
#                 1))

#cortex_biplot + theme(plot.margin = unit(c(3,3,3,3), "lines"))
#excavation_unit_biplot
#mass_biplot

#plot_grid(
#  plot_a, NULL, plot_b
#  , nrow = 1
#  , rel_widths = c(1, 0.05, 1)
#  , labels = c("A", "", "B")
#)

    #      axis = "none",
   #       scale = 1,
   #       nrow = 3,
    #      ncol = 1,
   #       rel_widths = c(1,1,1),
   #       rel_heights = c(1,1,1)

```

```{r cortex-biplot, fig.cap="PCA biplot for flake shape and dorsal cortex class."}
cortex_biplot
```

```{r excavation-unit-biplot, fig.cap="PCA biplot for flake shape and excavation unit."}
excavation_unit_biplot
```

```{r mass-biplot, fig.cap="PCA biplot for flake shape and mass cluster groups"}
mass_biplot
```

Figure \@ref(fig:cortex-biplot) shows flake shape (as defined within the first two principal components) in relation to dorsal cortex class. The primary class of dorsal cortex, representing initial, unretouched flakes in the reduction sequence, has the smallest range of shape variation. Secondary flakes, which have been retouched and thus have less dorsal cortex than the primary flakes, show a slightly larger range of shape variation. Lastly, the tertiary flakes at the end of the reduction sequence show the greatest amount of shape variation.

Shape variation is not equal along both principal components. For flakes in all three dorsal cortex classes, the largest differences between each inter-nested is along the first principal component. Slight differences are seen along the second principal component, with an increasing amount of variation as one goes from primary to secondary to tertiary flakes, but these changes are minor when compared to the larger changes seen in the first principal component.

Figure \@ref(fig:excavation-unit-biplot) shows the flake shapes projected on the first two principal components, and colored according to the excavation unit that they were recovered from. Excavation unit one is near the surface, and unit 13 is near the base. The general picture here is that shape variation changes little through the deposit. Units with fewer flakes, such as 13, show lower shape variation.

In Figure \@ref(fig:mass-biplot) we see that greater shape variation is found in the lightest mass classes. The two lightest mass clusters, one and two, are in purple and blue respectively. These first two mass classes span the largest area on the biplot, signifying a larger degree of variation overall.

```{r contribution-plots, fig.cap="Plots showing the contribution of landmarks to the overall shape variation among flakes."}
# let's look at all the PCs and how much each original # variable contributes to them

#view how much variation is described by each PC (ordered from PC1 to last PC)
# enframe(100*pca.geo$sdev^2/sum(pca.geo$sdev^2))

#get meaningful PCs
# getMeaningfulPCs(pca.geo$sdev^2, n = length(lands_no_outliers_names))

#estimate contribution (%) of landmarks on PC1
PC1_contrib <- fviz_contrib(pca.geo, choice = "var", axes = 1)
# take a look at the plot
# PC1_contrib

LM_contrib_PC1 <- data.frame(LM = rep(NA, nrow(PC1_contrib$data)/2),
                             PC1_contrib = rep(NA, nrow(PC1_contrib$data)/2))
for (i in 1:(nrow(PC1_contrib$data)/2)) {
  LM_contrib_PC1$PC1_contrib[i] <- sum(PC1_contrib$data[(i*2-1):(i*2),]$contrib)
  LM_contrib_PC1$LM[i] <- as.character(i)
}

# estimate contribution (%) of landmarks on PC2
PC2_contrib <- fviz_contrib(pca.geo, choice = "var", axes = 2)
# take a look at the plot
# PC2_contrib

LM_contrib_PC2 <- data.frame(LM = rep(NA, nrow(PC2_contrib$data)/2),
                             PC2_contrib = rep(NA, nrow(PC2_contrib$data)/2))
for (i in 1:(nrow(PC2_contrib$data)/2)) {
  LM_contrib_PC2$PC2_contrib[i] <- sum(PC2_contrib$data[(i*2-1):(i*2),]$contrib)
  LM_contrib_PC2$LM[i] <- as.character(i)
}

# examine landmarks with major contribution (%) to PC1 and PC2
# important_LMs_PC1 <- LM_contrib_PC1[LM_contrib_PC1$PC1_contrib > 100/(nrow(PC1_contrib$data)/2),]
# important_LMs_PC1[order(decreasing = T, important_LMs_PC1$PC1_contrib),]
# important_LMs_PC2 <- LM_contrib_PC2[LM_contrib_PC2$PC2_contrib > 100/(nrow(PC2_contrib$data)/2),]
# here is a table that summarises the most important landmarks in the PCA
# important_LMs_PC2[order(decreasing = T, important_LMs_PC2$PC2_contrib),]

library(cowplot)

plot_grid(PC1_contrib +
          scale_x_discrete(labels = 
                           as.factor(c("X Value: Top Left",
                          "Y Value: Top Left",
                          "X Value: Mid-Upper Left",
                          "Y Value: Mid-Upper Left",
                          "X Value: Mid-Lower Left",
                          "Y Value: Mid-Lower Left",
                          "X Value: Bottom Left",
                          "Y Value: Bottom Left",
                          "X Value: Distal Tip",
                          "Y Value: Distal Tip",
                          "X Value: Bottom Right",
                          "Y Value: Bottom Right",
                          "X Value: Mid-Lower Right",
                          "Y Value: Mid-Lower Right",
                          "X Value: Mid-Upper Right",
                          "Y Value: Mid-Upper Right",
                          "X Value: Top Right",
                          "Y Value: Top Right")))+
             theme(axis.text = element_text(size = 5)),
          PC2_contrib + 
          scale_x_discrete(labels = 
                           as.factor(c("X Value: Top Left",
                           "Y Value: Top Left",
                           "X Value: Mid-Upper Left",
                           "Y Value: Mid-Upper Left",
                           "X Value: Mid-Lower Left",
                           "Y Value: Mid-Lower Left",
                           "X Value: Bottom Left",
                           "Y Value: Bottom Left",
                           "X Value: Distal Tip",
                           "Y Value: Distal Tip",
                           "X Value: Bottom Right",
                           "Y Value: Bottom Right",
                           "X Value: Mid-Lower Right",
                           "Y Value: Mid-Lower Right",
                           "X Value: Mid-Upper Right",
                           "Y Value: Mid-Upper Right",
                           "X Value: Top Right",
                           "Y Value: Top Right")))+
             theme(axis.text = element_text(size = 5))
  )

```

Figure \@ref(fig:contribution-plots) shows the relative contribution of each of the original 18 variables put through our PCA. Looking at the supplementary table, we can see that the first principal component is primarily composed of the x-value for the distal tip landmark, with smaller contributions from the y-values from the two upper-most and mid-upper points. The second primary dimension is composed largely of the y-values for bottom-most and upper-most points.

# Discussion
  <!-- This portion partially consists of material from the ISKM presentation. Might want to reword things even more. Moreover, this section needs citations, which will bolster the overall content of this section. -->

  Our data visualizations reveal some noteworthy trends regarding the degree of core reduction. Figure \@ref(fig:excavation-unit-biplot) shows very little variation over time, with lots of overlap between groups in the biplot. Units with fewer flakes, such as 13, show lower shape variation. This generally confirms our conclusion from the radiocarbon ages that the deposit accumulated relatively quickly, and the assemblage represents knapping events occurring at a similar time with little, if any perceptible change over time. By contrast, a more definitive difference between groups in seen in Figure \@ref(fig:mass-biplot), where there is a high degree of shape variation in the lightest flakes. This reflects the production of different types and shapes of cores as the flakes become smaller and thus lighter. This is further supported by Figure \@ref(fig:cortex-biplot), where tertiary flakes with little dorsal cortex show the highest shape variation. This supports our interpretation of the previous result, with flakes from later in the reduction process showing greater shape variation as different types of cores were produced at the site.

 <!-- QH: As I write this I notice a major issue with the axes labels on the contribution plots. Instead of substituting the original labels on a one to one basis with our English labels, our plots are just printing out our list of English labels in the order they are listed in our code. This has to be fixed. I then need to redo my description of the plots, since it is all wrong and based off of the eroneous labels. -->

Looking at \@ref(fig:contribution-plots), the landmarks which contributed the most to each principal component were all located on the left side of our artefacts. This means that linking our initial flake measurements (maximum length, top width etc.) to either of the first two principal components is nearly impossible, as every one of our initial flake measurements (with the)

one can see that there is no consistent pattern in which certain initial flake measurements  align closely with either the first or second principal components. Interestingly, the landmarks which contributed the most to each principal component were all located on the left side of our artefacts..

The lack of distinct morphological groupings and the increased variability of shape in the smaller, tertiary flakes is consistent with previous studies. Other studies which investigated the morphology of flake outlines from South East Asia concluded that there is no significant relationship between flake outline and flake use (Borel et al., 2013; Borel et al., 2017). While our analysis not rely upon outlines, both outlines and landmarks are ultimately methods for measuring the morphology of flaked stone tools, and as such, the same models which explain outline variability can be cautiously applied to our analysis. The large amount of variability in flake shape, especially in the tertiary flakes, suggests that the knappers were not looking to create a specific shape of flake, with the amount of varitiblity increasing as flakes are struck off of increasingly differed cores. This behavior is consistent with processes of expedient knapping, where a large number of flakes are created in a relatively free-form manner and useful flakes are picked up and used after knapping is finished (Holdaway & Douglass, 2012; McCall, 2012). Processes of expedient knapping suggest that stone tools were of secondary importance in the material culture of the knappers (Sillitoe & Hardy, 2003). Elsewhere it has been suggested stone tools were primarily used to create bamboo tools in South East Asia (Mijares, 2001). A reliance upon bamboo tools, with stone flakes merely being an intermediary to produce bamboo tools, would explain the great variability in the lithic assemblages at Mau A.

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

## References

Archer, W., Djakovic, I., Brenet, M., Bourguignon, L., Presnyakova, D., Schlager, S., Soressi, M., & McPherron, S. P. (2021). Quantifying differences in hominin flaking technologies with 3D shape analysis. Journal of Human Evolution, 150, 102912. https://doi.org/10.1016/j.jhevol.2020.102912

Borel, Antony, Richard Cornette, and Michel Baylac. 2017. “Stone Tool Forms and Functions: A Morphometric Analysis of Modern Humans’ Stone Tools From Song Terus Cave (Java, Indonesia).” Archaeometry 59 (3): 455–71. https://doi.org/10.1111/arcm.12264.

Borel, Antony, Claire Gaillard, Marie-Hélène Moncel, Robert Sala, Emmanuelle Pouydebat, Truman Simanjuntak, and François Sémah. 2013. “How to Interpret Informal Flakes Assemblages? Integrating Morphological Description, Usewear and Morphometric Analysis Gave Better Understanding of the Behaviors of Anatomically Modern Human from Song Terus (Indonesia).” Journal of Anthropological Archaeology 32 (4): 630–46. https://doi.org/10.1016/j.jaa.2013.03.002.

Bradbury, Andrew P., and Philip J. Carr. "Flake Typologies and Alternative Approaches: An Experimental Assessment." Lithic Technology 20, no. 2 (1995): 100-15. Accessed June 11, 2021. http://www.jstor.org/stable/23273168.

Hoggard, C. S., McNabb, J., & Cole, J. N. (2019). The Application of Elliptic Fourier Analysis in Understanding Biface Shape and Symmetry Through the British Acheulean. Journal of Paleolithic Archaeology, 2(2), 115–133. https://doi.org/10.1007/s41982-019-00024-6

Holdaway, S., & Douglass, M. (2012). A Twenty-First Century Archaeology of Stone Artifacts. Journal of Archaeological Method and Theory, 19(1), 101–131. https://doi.org/10.1007/s10816-011-9103-6

McAdams, C., Morley, M. W., Fu, X., Kandyba, A. V., Derevianko, A. P., Nguyen, D. T., Doi, N. G., & Roberts, R. G. (2020). The Pleistocene geoarchaeology and geochronology of Con Moong Cave, North Vietnam: Site formation processes and hominin activity in the humid tropics. Geoarchaeology, 35(1), 72–97. https://doi.org/10.1002/gea.21758

McCall, G. S. (2012). Ethnoarchaeology and the Organization of Lithic Technology. Journal of Archaeological Research, 20(2), 157–203. https://doi.org/10.1007/s10814-011-9056-z

Mijares, A. S. B. (2001). An Expedient Lithic Technology in Northern Luzon (Philippines). Lithic Technology, 26(2), 138–152. https://doi.org/10.1080/01977261.2001.11720983

Mijares, A. 2008. “The late pleistocene to early holocene foragers of northern Luzon.”		Bulletin of the Indo-Pacific Prehistory Association 28: 99-107.

Okumura, M., & Araujo, A. G. M. (2014). Long-term cultural stability in hunter–gatherers: A case study using traditional and geometric morphometric analysis of lithic stemmed bifacial points from Southern Brazil. Journal of Archaeological Science, 45, 59–71. https://doi.org/10.1016/j.jas.2014.02.009

Radinović, M., & Kajtez, I. (2021). Outlining the knapping techniques: Assessment of the shape and regularity of prismatic blades using elliptic Fourier analysis. Journal of Archaeological Science: Reports, 38, 103079. https://doi.org/10.1016/j.jasrep.2021.103079

Riede, F., & Pedersen, J. B. (2018). Late Glacial Human Dispersals in Northern Europe and Disequilibrium Dynamics. Human Ecology, 46(5), 621–632. https://doi.org/10.1007/s10745-017-9964-8

Sillitoe, P., & Hardy, K. (2003). Living Lithics: Ethnoarchaeology in Highland Papua New Guinea. Antiquity, 77(297), 555–566. https://doi.org/10.1017/S0003598X00092619

Song, M., & Zhong, H. (2020). Efficient weighted univariate clustering maps outstanding dysregulated genomic zones in human cancers. Bioinformatics, 36(20), 5027–5036. https://doi.org/10.1093/bioinformatics/btaa613

Theska, T., Sieriebriennikov, B., Wighard, S. S., Werner, M. S., & Sommer, R. J. (2020). Geometric morphometrics of microscopic animals as exemplified by model nematodes. Nature Protocols, 15(8), 2611–2644. https://doi.org/10.1038/s41596-020-0347-z

Wang, H., & Song, M. (2011). Ckmeans.1d.dp: Optimal k-means Clustering in One Dimension by Dynamic Programming. The R Journal, 3(2), 29–33.

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```